---
title: "Ticket to Ride"
output: html_notebook
---

# Description

Analyse/play with Ticket to Ride map data.

# Setup

```{r, echo=FALSE}
rm(list=ls())
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(purrr)
library(igraph)
library(ggraph)
source("ttr-sparql.R")

theme_set(theme_minimal())
```


# Import data

```{r}
# Names of all routes from SPARQL
read_routes_sparql <- function(query) {
  qs(query) %>%
    select(-id)
}

# Read routes from CSV file
read_routes_csv <- function (csv_file) {
  read_csv(csv_file)
}
  
# routes <- read_from_sparql(q_all)
routes <- read_routes_csv("ttr-europe.csv") %>%
    mutate(colour = as.factor(colour))
  
glimpse(routes)
```

# Data analysis

## Route lengths

### Distribution of route lengths

```{r}
routes %>%
  group_by(length) %>%
  summarise(number = n())
```

### Name the longest routes

```{r}
routes %>%
  filter(length >= 6) %>%
  arrange(desc(length)) %>%
  select(length, name1, name2)
```


### Routes requiring locomotives

```{r}
routes %>%
  filter(locos > 0) %>%
  select(name1, name2, length, locos) %>%
  arrange(desc(locos))
```

### Tunnels

```{r}
routes %>%
  filter(isTunnel == TRUE) %>%
  select(name1, name2, length) %>%
  arrange(desc(length))
```


# Graph analysis

## Convert to a graph

```{r}
cities <- c(routes$name1, routes$name2) %>% unique %>% sort

g <- routes %>% 
  mutate(weight = length) %>%
  graph_from_data_frame(vertices = cities, directed = FALSE)

g %>% gorder()
```

## Visualise the whole map

```{r}
theme_no_axis <- theme(
    legend.position = "none", 
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank())
```

```{r}
g %>%
  ggraph(layout = 'igraph', algorithm = 'kk') +
  geom_edge_fan(
    aes(colour=colour, label=weight),
    width = 1.5, spread = 3, 
    label_size = 3, label_colour = "#99cccc") +
  geom_node_text(
    aes(label = name), 
    colour = "white", size = 3, 
    nudge_x = 0, nudge_y = 0.25) +
  geom_node_point(
    size = 2, colour = "#99ffff") +
  scale_edge_color_manual(
    values = c("black", "#9999ff", "green", "#999999", "orange", 
               "pink", "red", "#eeeeee", "yellow")) +
  theme(
    panel.background = element_rect(fill = "#333399")) +
  theme_no_axis +
  labs(
    title = "TTR Europe")
```


## Analysis of edges

Aggregate view of routes by colour

```{r}
routes %>%
  group_by(colour) %>%
  summarise(TotalLength = sum(length), Tunnels = sum(isTunnel), Locos = sum(locos))
```

## Graph metrics

Degree of each city

```{r}
num_neighbours <- function (g, v) {
  neighbors(g, v)$name %>% unique() %>% length()
}
```

Number of neighbouring vertices

```{r}
degree_n <- 
  data_frame(city = g %>% V %>% names) %>% 
  rowwise() %>%
  mutate(neighbours = num_neighbours(g, city))
```


```{r}
city_degree <- 
  g %>% 
  degree() %>% 
  data_frame(city = names(.), degree = .) %>%
  mutate(degree = as.integer(degree)) %>%
  left_join(degree_n, by = "city") %>%
  arrange(desc(degree))

```

Those with fewer neighbours than routes

```{r}
city_degree %>%
  filter(neighbours < degree)
```


Unique edges attached to the city with the highest degree

```{r}
g %>% incident(head(city_degree)$city) %>% attr("vnames") %>% unique()
```


Diameter and girth

```{r}
farthest_vertices(g)
message("----")
girth(g)
message("----")
mean_distance(g)
```

### Shortest path

```{r}
shortest_path <- function (g, from, to) {
  shortest_paths(g, from, to)$vpath[[1]]
}
```


### Minimum spanning tree

```{r}
gmst <- mst(g)
```

```{r}
gmst %>%
  ggraph(layout = 'igraph', algorithm = "kk") +
  geom_edge_fan(
    aes(colour=colour),
    width = 1.5, spread = 0) +
  geom_node_text(
    aes(label = name), 
    colour = "white", size = 3, 
    nudge_x = 0, nudge_y = 0.2) +
  geom_node_point(
    size = 1, colour = "#99ffff") +
  scale_edge_color_manual(
    values = c("black", "#9999ff", "green", "#999999", "orange", 
               "pink", "red", "#eeeeee", "yellow")) +
  theme(
    panel.background = element_rect(fill = "#333399")) +
  theme_no_axis +
  labs(
    title = "TTR Europe: minimal spanning tree")
```


# Export

Export routes to CSV

```{r, eval=FALSE}
routes %>%
  select(-id) %>%
  write_csv("ttr-europe.csv")
```

Export graph to GRAPHML

```{r, eval=FALSE}
g %>%
  write_graph(file = "ttr-europe.graphml", format = "graphml")
```

