---
title: "Ticket to Ride"
output: html_notebook
---

# Description

Analyse/play with TTR data retrieved over SPARQL.

# Setup

```{r, echo=FALSE}
rm(list=ls())
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(purrr)
library(igraph)
library(ggraph)
library(SPARQL)

theme_set(theme_minimal())
```

## Set up namespace prefixes and SPARQL endpoint

```{r}
prefixes = c(
  'rdf',  'http://www.w3.org/1999/02/22-rdf-syntax-ns#',
  'rdfs', 'http://www.w3.org/2000/01/rdf-schema#',
  'xsd',  'http://www.w3.org/2001/XMLSchema#',
  'skos', 'http://www.w3.org/2004/02/skos/core#',
  'ttr',  'http://alphajuliet.com/ns/ttr#')

endpoint = 'https://dydra.com/alphajuliet/ttr-europe/sparql'

q_sparql <- partial(SPARQL, url = endpoint, ns = prefixes, format = "xml")
qs <- function(q) {
  q_sparql(query = q)$results %>% as.tibble()
}
```

# SPARQL Queries

```{r}
#----------------------
# First 20 triples in the repo

q_first20 = "
select * where {
  ?s ?p ?o
} limit 20"

#----------------------
# Return the distributions by colour

q_sbc = "
prefix : <http://alphajuliet.com/ns/ttr#>
select ?colour (count(?r) as ?routes) (sum(?len) as ?length)
where {
  ?r a :Route ;
     :colour ?c ;
     :length ?len .
  ?c rdfs:label ?colour .
} 
group by ?colour
order by desc(?length)"

#----------------------
# Return the distributions by length

q_dlen = "
prefix : <http://alphajuliet.com/ns/ttr#>
select ?len (count(?r) as ?number)
where {
  ?r a :Route ;
     :length ?len .
} 
group by ?len
order by ?len"

#----------------------
# Return all route data

q_all = "
prefix : <http://alphajuliet.com/ns/ttr#>
select distinct ?id ?name1 ?name2 ?colour ?length ?isTunnel ?locos
where {
  ?id a :Route ;
     :colour ?clr ;
     :length ?length ;
     :isTunnel ?isTunnel ;
     :locomotives ?locos ;
     :end ?city1 , ?city2 .
  ?city1 rdfs:label ?name1 .
  ?city2 rdfs:label ?name2 .
  ?clr rdfs:label ?colour .
  filter (?city1 < ?city2)
}"
```


# Import data

```{r}
# Names of all routes
routes <- qs(q_all) %>%
  mutate(colour = as.factor(colour))
glimpse(routes)
```

# Data analysis

## Route lengths

### Distribution of route lengths

```{r}
routes %>%
  group_by(length) %>%
  summarise(number = n()) %>%
  ggplot(aes(x = length, y = number)) +
  geom_col(fill = "#99aaff") + 
  geom_text(aes(label = number), nudge_y = 2) +
  theme_dark()
```

### Longest routes

```{r}
routes %>%
  filter(length >= 6) %>%
  arrange(desc(length)) %>%
  select(length, name1, name2)
```


### Routes requiring locomotives

```{r}
routes %>%
  filter(locos > 0) %>%
  select(name1, name2, length, locos) %>%
  arrange(desc(locos))
```

### Tunnels

```{r}
routes %>%
  filter(isTunnel == TRUE) %>%
  select(name1, name2, length) %>%
  arrange(desc(length))
```


# Graph analysis

## Convert to a graph

```{r}
cities <- c(routes$name1, routes$name2) %>% unique %>% sort

g <- routes %>% 
  select(-id) %>%
  mutate(weight = length) %>%
  graph_from_data_frame(vertices = cities, directed = FALSE)

g %>% gorder()
```

## Visualise the whole map

```{r}
theme_no_axis <- theme(
    legend.position = "none", 
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank())
```

```{r}
g %>%
  ggraph(layout = 'igraph', algorithm = 'kk') +
  geom_edge_fan(
    aes(colour=colour, label=weight),
    width = 1.5, spread = 3, 
    label_size = 3, label_colour = "#99cccc") +
  geom_node_text(
    aes(label = name), 
    colour = "white", size = 3, 
    nudge_x = 0, nudge_y = 0.25) +
  geom_node_point(
    size = 2, colour = "#99ffff") +
  scale_edge_color_manual(
    values = c("black", "#9999ff", "green", "#999999", "orange", 
               "pink", "red", "#eeeeee", "yellow")) +
  theme(
    panel.background = element_rect(fill = "#333399")) +
  theme_no_axis +
  labs(
    title = "TTR Europe")
```


## Analysis of edges

Aggregate view of routes by colour

```{r}
routes %>%
  group_by(colour) %>%
  summarise(TotalLength = sum(length), Tunnels = sum(isTunnel), Locos = sum(locos))
```

## Graph metrics

Degree of each city

```{r}
city_degree <- 
  g %>% 
  degree %>% 
  data_frame(city = names(.), degree = .) %>%
  mutate(degree = as.integer(degree)) %>%
  arrange(desc(degree))
```

```{r}
city_degree %>%
  ggplot(aes(reorder(city, degree), degree)) +
  geom_col(width = 0.3, fill = "red") +
  scale_y_continuous(breaks = seq(0, 10, 1)) +
  coord_flip() + 
  theme_minimal() +
  theme(axis.text.y = element_text(size = 6)) +
  labs(
    title = "Degree of cities",
    x = "City",
    y = "degree")
```

Edges attached to the city with the highest degree

```{r}
g %>% incident(head(city_degree)$city) %>% attr("vnames")
```


Diameter and girth

```{r}
farthest_vertices(g)
message("----")
girth(g)
message("----")
mean_distance(g)
```

### Minimum spanning tree

```{r}
gmst <- mst(g)
```

```{r}
gmst %>%
  ggraph(layout = 'igraph', algorithm = "kk") +
  geom_edge_fan(
    aes(colour=colour),
    width = 1.5, spread = 0) +
  geom_node_text(
    aes(label = name), 
    colour = "white", size = 3, 
    nudge_x = 0, nudge_y = 0.2) +
  geom_node_point(
    size = 1, colour = "#99ffff") +
  scale_edge_color_manual(
    values = c("black", "#9999ff", "green", "#999999", "orange", 
               "pink", "red", "#eeeeee", "yellow")) +
  theme(
    panel.background = element_rect(fill = "#333399")) +
  theme_no_axis +
  labs(
    title = "TTR Europe: minimal spanning tree")
```


# Export

Export routes to CSV

```{r, eval=FALSE}
routes %>%
  select(-id) %>%
  write_csv("ttr-europe.csv")
```

Export graph to GRAPHML

```{r, eval=FALSE}
g %>%
  write_graph(file = "ttr-europe.graphml", format = "graphml")
```


# Simulation

```{r}

```

